---
title: "DS2_HW5"
author: "Siyan Chen"
date: "4/24/2019"
output: pdf_document
---

```{r setup, include=FALSE}
library(ISLR)
library(caret)
data(OJ)
head(OJ)
```


### (a) Fit a support vector classifier (linear kernel) to the training data with Purchase as the response and the other variables as predictors. What are the training and test error rates?

```{r}
set.seed(1)
rowtrain = createDataPartition(y = OJ$Purchase,
                               p = 800/1070,
                               list = FALSE)

ctr1 = trainControl(method = "cv")

set.seed(1)
svml.fit = train(Purchase~.,
                data = OJ[rowtrain,],
                method = "svmLinear2",
                preProcess = c("center", "scale"),
                tuneGrid = data.frame(cost = exp(seq(-5,1,len = 20))),
                trControl = ctr1)
ggplot(svml.fit, highlight = TRUE)
best.svml = svml.fit$finalModel
summary(best.svml)

### train error
pred.svml_train = predict(svml.fit, newdata = OJ[rowtrain,])
1-confusionMatrix(data = pred.svml_train, reference = OJ$Purchase[rowtrain])$overall[["Accuracy"]]


### test error
pred.svml = predict(svml.fit, newdata = OJ[-rowtrain,])
mean(pred.svml != OJ[-rowtrain,]$Purchase)
# OR 1-confusionMatrix(data = pred.svml, reference = OJ$Purchase[-rowtrain])$overall[["Accuracy"]]
```

The trainning error rate is 0.1560549 and test error rate is 0.1858736.

### (b) Fit a support vector machine with a radial kernel to the training data. What are the training and test error rates?

```{r}
svmr.grid = expand.grid(C = exp(seq(-3, 3, len = 10)),
                        sigma = exp(seq(-8, -3, len = 5)))
set.seed(1)
svmr.fit = train(Purchase~.,
                 data = OJ[rowtrain,],
                 method = "svmRadial",
                 preProcess = c("center", "scale"),
                 tuneGrid = svmr.grid,
                 trControl = ctr1)

ggplot(svmr.fit, highlight = TRUE)

### train error
pred.svmr_train = predict(svmr.fit, newdata = OJ[rowtrain,])
1-confusionMatrix(data = pred.svmr_train, reference = OJ$Purchase[rowtrain])$overall[["Accuracy"]]


### test error rate
pred.svmr = predict(svmr.fit, newdata = OJ[-rowtrain,])
mean(pred.svmr != OJ[-rowtrain,]$Purchase)
```


Train error rate is  0.1510612 and test error rate is 0.2007435.

### (c) Which approach seems to give a better result on this data?

```{r}
resamp = resamples(list(svml = svml.fit,svmr = svmr.fit))
bwplot(resamp)
summary(resamp)
```

According the boxplot, both accuracy and kappa of the model with radial kernel is greater which suggests that it has smaller train error rate and greater inter-rater aggrement. Therefore, model with radial kernel is better.
