---
title: "hw3_bm"
author: "Siyan Chen"
output:
  pdf_document: default
  html_document: default
---

# Problem 1

```{r setup, include=FALSE}
age = c(25, 35, 45, 55, 65, 75, 25, 35, 45, 55, 65, 75 )
al_consumption = c(0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1) #0:<80g, #1: >80g  
y= c(0, 5, 21, 34, 36, 8, 1, 4, 25, 42, 19, 5)
m = c( 106, 169, 159, 173, 124, 39, 10, 30, 54, 69, 37, 5)
df = data.frame(age, al_consumption, y, m)
```

### fit model 

In this question, data follows binomail distribution y ~ bin(n, m). Alchol consumption are categorical variables. Daily alchol consumption less than 80g are set to reference level. 
Based on the results, model is fitted to be:
$log\frac{\pi}{1-\pi}$ = -5.023449 + 0.061579*age + 1.78*I{Daily alcohol consumption>80g}


```{r, include = FALSE}
logit.prosp = glm(cbind(y, m-y) ~ age + al_consumption, family = binomial(link = 'logit'))

summary(logit.prosp) 
```


## 1.1

Since we fit the prospective model to the retrospective data, intercept $\beta_0$ is not meaningful to interpret.

Based on the r ouput, log odds of getting esophageal cancer increases 0.061579 for 1 years older if daily alcohol consumption is constant. The log odds ratio of getting esophageal cancer is 1.78 for individuals with > 80g daily alcohol consumption versus individuals with < 80g dailyb alcohol consumtion when other variables are constant. 





# Problem 2

The reference level is seed O.aegytiaca 75 in bean media. 

Model: $log\frac{\pi}{1-\pi}$ = -0.43-0.2705*I{seed 73}+1.0647*I{cucumber media} 


```{r, include=FALSE}
### data input by grouped data
seed = as.factor(c("A","A","B", "B"))
median =as.factor(c ("bean","cucu","bean","cucu"))
y = c(99, 201, 49, 75)
m = c(272, 295, 123, 141)
data2 = data.frame(seed, median, y, m)
```



```{r, include=FALSE}
### model
logit_model =  glm(cbind(y, m-y) ~ seed + median, family = binomial(link = 'logit'))

summary(logit_model)
```
### 2.1
Log odds of germination is -0.43 for seed of O.aegytiaca 75 in the bean median. 

Log odds ratio of germination for seed of O.aegytiaca 75 versus seed of O.aegytiaca 73 is -0.27 in the bean median. 

Log odds ratio of germination in bean median versus cucumber media is 1.06 for seed of O.aegytiaca 75.

```{r, include=FALSE}
dev = sum(residuals(logit_model,type='deviance')^2)
1-pchisq(dev, 1)  # p value is significant, so we reject the hypothesis, so model does not fit well.

G.stat=sum(residuals(logit_model,type='pearson')^2)# pearson chisq 
phi=G.stat/(4-3)
```

# test over-dispersion (half normal plot)

```{r, include= FALSE}
res=residuals(logit_model,type='pearson')
```

```{r, echo = FALSE}
plot(qnorm((4+1:4+0.5)/(2*4+1.125)),sort(abs(res)),xlab='Expected Half-Normal Order Stats',ylab='Ordered Abs Pearson Residuals', ylim=c(0,2), xlim=c(0,2))
abline(a=0,b=1)
abline(a=0,b=sqrt(phi),lty=2)

```

```{r, include=FALSE}
summary(logit_model,dispersion=phi)
```

### 2.2
Deviance is used for testing goodness of fit. P value is 0.011 which is small than significant level.Therefore, this model does not fit well.

There is over disoersion based on the plot and the estimate of dispersion parameter is 6.48. After updating model, we get same coefficient but different variance for coefficient. 

### 2.3
Possible reseason can be that correlation exists between each seed germination. 















#CODE


```{r, results= "hide"}
age = c(25, 35, 45, 55, 65, 75, 25, 35, 45, 55, 65, 75 )
al_consumption = c(0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1) #0:<80g, #1: >80g  
y= c(0, 5, 21, 34, 36, 8, 1, 4, 25, 42, 19, 5)
m = c( 106, 169, 159, 173, 124, 39, 10, 30, 54, 69, 37, 5)
df = data.frame(age, al_consumption, y, m)

logit.prosp = glm(cbind(y, m-y) ~ age + al_consumption, family = binomial(link = 'logit'))

summary(logit.prosp) 



```


```{r,  results= "hide"}
seed = as.factor(c("A","A","B", "B"))
median =as.factor(c ("bean","cucu","bean","cucu"))
y = c(99, 201, 49, 75)
m = c(272, 295, 123, 141)
data2 = data.frame(seed, median, y, m)

### model
logit_model =  glm(cbind(y, m-y) ~ seed + median, family = binomial(link = 'logit'))

summary(logit_model)

dev = sum(residuals(logit_model,type='deviance')^2)
1-pchisq(dev, 1)  # p value is significant, so we reject the hypothesis, so model does not fit well.

G.stat=sum(residuals(logit_model,type='pearson')^2)# pearson chisq 
phi=G.stat/(4-3)

res=residuals(logit_model,type='pearson')

plot(qnorm((4+1:4+0.5)/(2*4+1.125)),sort(abs(res)),xlab='Expected Half-Normal Order Stats',ylab='Ordered Abs Pearson Residuals', ylim=c(0,2), xlim=c(0,2))
abline(a=0,b=1)
abline(a=0,b=sqrt(phi),lty=2)

summary(logit_model,dispersion=phi)
```

