---
title: "BM_HW4"
author: "Siyan Chen_sc4456"
date: "3/4/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(nnet)
library(MASS)
library(scales)
library(tidyverse)
```

# data input


```{r, message=FALSE, warning=FALSE}
df = data.frame(ls_y = c(65, 130, 67, 34, 141, 130), ms_y = c(54, 76, 48, 47, 116, 105), hs_y = c(100, 111, 62, 100, 191, 104), contact = c(rep("low", 3), rep("high", 3)), type = c("tower_block", "apartment", "house", "tower_block", "apartment", "house"))
df$contact = factor(df$contact, level = c("low", "high"))
df$type = factor(df$type, level = c("tower_block", "apartment", "house"))
```

#1 Pair-wise Table

```{r,echo = FALSE}
df %>% 
  group_by(contact) %>% 
  summarise(ls = sum(ls_y), ms = sum(ms_y), hs = sum(hs_y)) %>% 
  mutate(ls_p = ls/(ls + ms + hs), ms_p = ms/(ls + ms + hs), hs_p = hs/(ls + ms + hs)) %>% 
  mutate(ls_p = percent(ls_p),
         ms_p = percent(ms_p),
         hs_p = percent(hs_p)) %>% 
  knitr::kable()

df %>% group_by(type) %>% 
  summarise(ls = sum(ls_y), ms = sum(ms_y), hs = sum(hs_y)) %>% 
  mutate(ls_p = ls/(ls + ms + hs), ms_p = ms/(ls + ms + hs), hs_p = hs/(ls + ms + hs)) %>% 
  mutate(ls_p = percent(ls_p),
         ms_p = percent(ms_p),
         hs_p = percent(hs_p)) %>% 
  knitr::kable()

df %>% 
  mutate(ls_p = percent(ls_y/(ls_y+ms_y+hs_y)), ms_p = percent(ms_y/(ls_y+ms_y+hs_y)), hs_p = percent(hs_y/(ls_y+ms_y+hs_y))) %>% 
  knitr::kable()
  
```

According to the tables, when residents have high degree contact, the proportion of response in medium or high satisfaction categories tend to increase. When residents lives in tower block, the proportion of response in high satisfaction categorys tend to increase. When residents lives in tower block and have high degree contact, the proportion of response in high satisfaction category
is greatest (55.2%).

# 2 Nomial logisitic regression

```{r, message=FALSE, warning=FALSE}
multi_model = multinom(cbind(ls_y, ms_y, hs_y) ~ contact + type, data = df)
summary(multi_model)
```

Interpretation:
 $log \frac{\pi_{medium}}{\pi_{low}} = \beta_{10} + \beta_{11}*I\{contact = high\} + \beta_{12}*I\{type = apartment\} +\beta_{13}*I\{type = house\}$

$\beta_{10}$: Log odds for response of Medium satisfaction versus Low satisfcation is -0.107 if residents live in tower block and have low contacts with others.

$\beta_{11}$: Log odds for response of Medium satisfaction versus Low satisfcation increases 0.296 if residents have high contacts while house type does not changes.

$\beta_{12}$ ; Log odds for response of Medium satisfaction versus Low satisfcation decreases 0.407 if residents live in apartment while contact frequency does not change.

$\beta_{13}$Log odds for response of Medium satisfaction versus Low satisfcation decreases 0.337 if residents live in house while contact frequency does not change.

2. $log \frac{\pi_{high}}{\pi_{low}}=\beta_{20} + \beta_{21}*I\{contact = high\} + \beta_{22}*I\{type = apartment\} +\beta_{23}*I\{type = house\}$

$\beta_{20}$:Log odds for response of High satisfaction versus Medium satisfcation is 0.561 if residents live in tower block and have low degree of contacts with others.

$\beta_{21}$:Log odds for response of High satisfaction versus Medium satisfcation increases 0.328 if residents have high degree of contacts while house type does not changes.

$\beta_{22}$:Log odds for response of High satisfaction versus Medium satisfcation decreases 0.642 if residents live in apartment while degree of contact does not change.

$\beta_{23}$:Log odds for response of High satisfaction versus Medium satisfcation decreases 0.946 if residents live in house while degree of contact does not change


# 95% CI

```{r, results = FALSE}
beta = summary(multi_model)$coefficients
se1=sqrt(vcov(multi_model)[1,1])
se2 = sqrt(vcov(multi_model)[2,2])
se3=sqrt(vcov(multi_model)[3,3])
se4=sqrt(vcov(multi_model)[4,4])
se5=sqrt(vcov(multi_model)[5,5])
se6=sqrt(vcov(multi_model)[6,6])
se7=sqrt(vcov(multi_model)[7,7])
se8=sqrt(vcov(multi_model)[8,8])
exp(beta[1]+c(qnorm(0.025),0,-qnorm(0.025))*se1)
exp(beta[2]+c(qnorm(0.025),0,-qnorm(0.025))*se5)
exp(beta[3]+c(qnorm(0.025),0,-qnorm(0.025))*se2)
exp(beta[4]+c(qnorm(0.025),0,-qnorm(0.025))*se6)
exp(beta[5]+c(qnorm(0.025),0,-qnorm(0.025))*se3)
exp(beta[6]+c(qnorm(0.025),0,-qnorm(0.025))*se7)
exp(beta[7]+c(qnorm(0.025),0,-qnorm(0.025))*se4)
exp(beta[8]+c(qnorm(0.025),0,-qnorm(0.025))*se8)

```

 model1:
 CI $\beta_{10}$: [0.6663249 0.8982882 1.2110033]\
    $\beta_{11}$:  [1.041831 1.344444 1.734954]\
    $\beta_{12}$:  [0.4759238 0.6658082 0.9314528]\
    $\beta_{13}$:  [0.5012894 0.7138538 1.0165531]\


We are 95% confident that odds for response of Medium satisfaction versus Low satisfcation is between 0.666 and 1.211 if residents live in tower block and have low contacts with others.

We are 95% confident that odds ratio for response of Medium satisfaction versus Low satisfcation for high contact vs low contact is between 1.042 and 1.735.

We are 95% confident that odds ratio for response of Medium satisfaction versus Low satisfcation for apartment vs tower_block is between 0.476 and 0.931.

We are 95% confident that odds ratio for response of Medium satisfaction versus Low satisfcation for house vs tower_block is between 0.501 and 1.017.


Simialr explaination for model 2

model2:
 CI $\beta_{20}$: [1.350177 1.752027 2.273480]\
    $\beta_{21}$: [1.101402 1.388503 1.750442]\
    $\beta_{22}$: [0.3922944 0.5264512 0.7064869]\
    $\beta_{23}$: [0.2813932 0.3884396 0.5362080]\
    
We are 95% confident that odds for response of Medium satisfaction versus High satisfcation is between 1.35 and 2.27 if residents live in tower block and have low contacts with others.


# goodness of fit

```{r}
pihat = predict(multi_model, type = "probs")
m = rowSums(df[,1:3])
res_pearson  = (df[,1:3] - pihat*m)/sqrt(pihat*m) ## pearson residuals 
G.stats = sum(res_pearson^2)
G.stats
pval = 1-pchisq(G.stats, df = (3-1)*(6-3))
pval # could not reject the null hypothesis, fit well
```

The pearson deiviance of model is 6.932334, which follows chi_square distribution with df = (3-1)*(6-3).
According p value, this model fits data well.

# 3 Proportion Model

```{r, message=FALSE, warning=FALSE}
freq = c(df$ls_y, df$ms_y, df$hs_y)
res = c(rep(c("ls", "ms", "hs"), c(6, 6, 6)))
res = factor(res, level = c("ls", "ms", "hs"), ordered = T)
df2 = data.frame(res, contact = c("low", "low", "low", "high","high","high"), type = c(rep(c("tower_block", "apartment", "house"),c(3))), freq )
df2$type = factor(df2$type, level = c("tower_block", "apartment", "house" ))
df2$contact = factor(df2$contact, level = c("low","high"))

```

```{r, message=FALSE, warning=FALSE}
# fit proportional odds model
polr_model = polr(res ~ contact + type, data = df2, weights = freq)
summary(polr_model)
```

$log\frac{P(Y\le j)}{P(Y>j)} = aj + 0.2524*I\{contact = high\}-0.5009*I\{type = apartment\}-0.7362*I\{type = house\}$

ls|ms -0.9973 is a1
ms|hs  0.1152 is a2

aj: the log odds of falling into or below category j when contact = low and type = tower block

a2: The estimated log odds of response falls in medium satisfaction or lower is 0.1152 when contact = low and type = tower block

$\beta_k$: incrrease in log-odds of falling into or below category with one unit increase in $x_k$ holding all other x vaiables constant.

$\beta_1 = 0.2524$: Increase in log-odds of falling into or below category when contact changes to high while other x variables constant.




#4 

```{r, message=FALSE, warning=FALSE}
pihat1 = predict(polr_model,df,type='p')
res.pearson=(df[,1:3] - pihat1*m)/sqrt(pihat1*m)
G=sum(res.pearson^2)

p_polr = predict(polr_model, df, type = 'p')
res.pearson_2 = (df[,1:3] - p_polr*m)/sqrt(p_polr*m)

abs(res.pearson_2)[6,3]

```

According to the pearson residuals for matrix, the greatest discrepancies between the observed frequencies and expected frequencies estimated from the model occurs in category of high frequency of contact and house type.



