---
title: "bm2_Hw5"
author: "Siyan Chen_sc4456"
date: "3/7/2019"
output:
  pdf_document: default
  html_document: default
---

```{r, include=FALSE}
library(tidyverse)
library(pscl)
```

# Problem 1

```{r setup, include=FALSE}
crab_df = read.delim("./data/HW5-crab.txt", header = TRUE, sep = "")
```

### a

```{r, warning=FALSE,message=FALSE}
m1 = glm(Sa~W, family = poisson, data = crab_df)
summary(m1)
```

```{r,warning=FALSE,message=FALSE}
#gof
res_pearson = residuals(m1, crab_df, type = "pearson")
G = sum(res_pearson^2)
pval = 1-pchisq(G, 171)
pval  # fit bad
```

$log\lambda=\beta_0+\beta_1*x$\
$log\lambda=-3.30+0.164*x$\
$\beta_0$:The log number of satellites for each female when carapace width is 0, which is not meaningful.\
$\beta1$:The log number of satellites for each female increases 0.164 for one unit increase in carapace width.

This model follows chi-square distribution with degree of 171. The p value is 0, which is significant. Therefore, this model is not good fit. 



### b

```{r, warning=FALSE,message=FALSE}
m2 = glm(Sa ~ W + Wt, family = poisson, data = crab_df)
summary(m2)

test.stat = m1$deviance - m2$deviance 
df = 1
pval = 1 - pchisq(test.stat, df = 1)
pval # reject the null hypothesis, go with the bigger model
```

Deviance follows chi-square distribution with degree of 1. P value is significant, so we reject the null hypothesis, go with the bigger model(M2).

### c

```{r, warning=FALSE,message=FALSE}
res_pearson2 = residuals(m2, crab_df, type = "pearson")
G2 = sum(res_pearson2^2)
1-pchisq(G2, 170) # not good fit
phi = G2/(170)

summary(m2, dispersion = phi)
```

```{r,echo=FALSE}
plot(qnorm((173+1:173+0.5)/(2*173+1.125)),sort(abs(res_pearson2)), xlab='Expected Half-Normal Order Stats') + abline(a=0,b=1) + abline(a=0,b=sqrt(phi),lty=2)
```


Based on the plot, there is over dispersion om M2. The estimate of disperion parameter is 3.16. 
After updating the model, we got model with same coefficients but different variance of each coefficient.

$log\lambda=\beta_0+\beta_1*x_1 + \beta_2*x_2$\
$log\lambda=-1.29+0.046*x_1 + 0.447*x_2$\
$\beta_0$:The log number of satellites for each female when carapace width(W) is 0 and weight(Wt) is 0, which is not meaningful.\
$\beta_1$:The log number of satellites for each female increases 0.046 for one unit increase in carapace width(W).\
$\beta_2$:The log number of satellites for each female increases 0.447 for one unit increase in weight(Wt).

# Problem 2

```{r}
par_df = read.delim("./data/HW5-parasite.txt", header = TRUE, sep = "") %>% 
  mutate(Year = as.factor(Year),
         Area = as.factor(Area))

```

### a

```{r}
p_model = glm(Intensity ~ I(Area) + I(Year) + Length, family = poisson, data = par_df)
summary(p_model)
```

$log(\lambda) = \beta_0+\beta_1*I(area2)+\beta_2*I(area3)+\beta_3*I(area4)+\beta_4*I(year2000)+\beta_5(year2001)$\

$\beta_0$ is not meaningful to interpret.\
$\beta_1$:Log intensity rate of parasite on fish in area 2 decreases 0.212 compared to area 1, when other variables are constant.\
$\beta_2$:Log intensity rate of parasite on fish in area 3 decreases 0.117 compared to area 1, when other variables are constant.\
$\beta_3$:Log intensity rate of parasite on fish in area 4 increases 1.405 compared to area 1, when other variables are constant.\
$\beta_4$:Log intensity rate of parasite on fish in year 2000 increases 0.670 compared to year 1999, when other variables are constant.\
$\beta_5$:Log intensity rate of parasite on fish in year 2001 decreases 0218 compared to year 1999, when other variables are constant.\
$\beta_6$:Log intensity rate of parasite on fish decreases 0.0284 for one unit increase in length when other variables are constant.

### b

```{r}
r = residuals(p_model, data = par_df, type = "pearson")
G_stats = sum(r^2)
pval = 1-pchisq(G_stats, 1184)
pval
```

The pearson residuals follows chi-square distribution with degree of 1184(G = 42164.97). P value is 0 which is significant. Therefore, we reject the null hypothesis and this model is not good fit. 

### c

```{r}
i_model = zeroinfl(Intensity ~ I(Area) + I(Year) + Length, data = par_df)
summary(i_model)

```

We use zero-inflated Possion model to classify the response at risk and response not at risk.
$Z_i$ is a latent binary variable that generates structural zeros.
$P(Z_i=0)=\pi_i$ \
The response satisfies:
$Y_i|(Z_i=0)=0$\
$Y_i|(Z_i=0)$~possion($\lambda_i$)\
Therefore, we get 
$log(\lambda_i)=\beta_0 +\beta_1*I(area 2)+\beta_2*I(area3)+\beta_3*I(area4)+\beta_4*I(year2000)+\beta_5(year2001)$ \
$log(\frac{\pi_i}{1-\pi_i})=z_i\gamma = \alpha_0 +\alpha_1*I(area 2)+\alpha_2*I(area3)+\alpha_3*I(area4)+\alpha_4*I(year2000)+\alpha_5(year2001)$ \

$\beta_0$ is not meaningful to interpret.\
$\beta_1$:Log intensity rate of parasite on fish in area 2 increases 0.27 compared to area 1, when other variables are constant.\
$\beta_2$:Log intensity rate of parasite on fish in area 3 increases 0.15 compared to area 1, when other variables are constant.\
$\beta_3$:Log intensity rate of parasite on fish in area 4 increases 0.94 compared to area 1, when other variables are constant.\
$\beta_4$:Log intensity rate of parasite on fish in year 2000 increases 0.29 compared to year 1999, when other variables are constant.\
$\beta_5$:Log intensity rate of parasite on fish in year 2001 decreases 0.045 compared to year 1999, when other variables are constant.\
$\beta_6$:Log intensity rate of parasite on fish decreases 0.037 for one unit increase in length when other variables are constant.\
$\alpha_0$:not meaningful for length = 0.\
$\alpha_1$: Log odd ratio of fish not at risk of parasite vs. at risk is 0.72 for fish in area 2 versus in area 1, holding other variables constant.\
$\alpha_2$:Log odd ratio of fish not at risk of parasite vs. at risk is 0.66 for fish in area 3 versus in area 1, holding other variables constant.\
$\alpha_3$:Log odd ratio of fish not at risk of parasite vs. at risk is -1.02 for fish in area 4 versus in area 1, holding other variables constant.\
$\alpha_4$:Log odd ratio of fish not at risk of parasite vs. at risk is -0.75 for fish in year 2000 versus in year 1999, holding other variables constant.\
$\alpha_5$:Log odd ratio of fish not at risk of parasite vs. at risk is 0.45 for fish in year 2001 versus in year 1999, holding other variables constant.\
$\alpha_6$:Log odd ratio of fish not at risk of parasite vs. at risk decreases 0.010 for one unit increas in length, holding other variables constant.\