---
title: "HW6"
author: "Siyan Chen"
date: "4/1/2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(tidyverse)
library(nlme)
dental_df = read.delim("/Users/siyanchen/Desktop/data/HW6-dental copy.txt", header = TRUE, sep = "") %>%
  janitor::clean_names()
  
```

# PROBLEM1

$var(Y_{ij})=var(\mu+b_i+e_{ij})$
$=var(b_i)+var(e_{ij})$
$={\sigma_b}^2+{\sigma_e}^2$

$cov(Y_{ij},Y_{ik}) = E[(Y_{ij}-\mu_{ij})(Y_{ik}-\mu_{ik})]$
$=E(b_i+e_{ij})(b_i+e_{ik})$
$=E({b_i}^2+e_{ij}*e_{ik}+b_i*e_{ik}+b_i*e_{ij})$
$=E({b_i}^2)+E(e_{ij}*e_{ik})+E(b_i*e_{ik})+E(b_i*e_{ij})$
$=E(b_i^2)$
$=(Eb_i)^2+var(b_i)$
$=\sigma_b^2$

$corr(Y_{ij},Y_{ik})=\frac{cov(Y_{ij},Y_{ik})}{\sqrt{ var(Y_{ij})*var(Y_{ik})}}$

$={\sigma_b}^2/({\sigma_b}^2+{\sigma_e}^2)$

# PROBLEM 2
### 2.1

```{r}
ggplot(dental_df, aes(x = age, y = distance, group = child)) + geom_line(aes(color = factor(gender))) 
```

### 2.2
$E(Y_{ij})=\beta_0+\beta_1*age_{ij}$\
$var(Y_{ij})=var(a_i)+var(b_k)+var(e_{ij})$
$={\sigma_a}^2+{\sigma_b}^2+{\sigma_e}^2$

### 2.3

```{r}
### compound symmetry
comsym = gls(distance ~ age + gender,dental_df, correlation=corCompSymm(form = ~ 1|child),weights=varIdent(form = ~ 1|age), method="REML")
summary(comsym)
corMatrix(comsym$modelStruct$corStruct)[[1]]
```


```{r}
### exponential
exp = gls(distance ~ age + gender,dental_df, correlation=corExp(form = ~ 1|child), method="REML")
summary(exp)
corMatrix(exp$modelStruct$corStruct)[[1]]
```


```{r}
### autoregressive
auto1 = gls(distance ~ age + gender,dental_df, correlation=corAR1(form = ~ 1|child), method="REML")
summary(auto1)
corMatrix(auto1$modelStruct$corStruct)[[1]]
```
The three method produce similar coefficient estimates. For model with compound symmetry covariance, variance is constant and correlation between any two visits is 0.617.
Exponential and autogressive covariance have same covariance and coefficient estimae.Exponential covariance is a gneneralization of AR(1) and correlation decreases exponentially as age difference increases. It reduces to AR(1) when all response time are same and the correlation decreases between two ages. 
